---
title: "Conditional Probability"
author: "Cameron Hayman"
date: "02/17/2025"

format: 
  html:  # You will quite likely want to change all but the last one, to taste
    theme: superhero  
    mainfont: monospace
    highlight-style: github
    title-block-banner: true
    embed-resources: true
---

**Abstract:**

This is a technical blog post of **both** an HTML file *and* [.qmd file](https://raw.githubusercontent.com/cd-public/D505/refs/heads/master/hws/src/cond.qmd) hosted on GitHub pages.

# 0. Quarto Type-setting

- This document is rendered with Quarto, and configured to embed an images using the `embed-resources` option in the header.
- If you wish to use a similar header, here's is the format specification for this document:

```email
format: 
  html:
    embed-resources: true
```

# 1. Setup

**Step Up Code:**

```{r}
sh <- suppressPackageStartupMessages
sh(library(tidyverse))
sh(library(caret))
wine <- readRDS(gzcon(url("https://github.com/cd-public/D505/raw/master/dat/pinot.rds")))
head(wine)
```

# 2. Conditional Probability

Calculate the probability that a Pinot comes from Burgundy given it has the word 'fruit' in the description.

$$
P({\rm Burgundy}~|~{\rm Fruit})
$$

```{r}
# TODO
wine_1 <- wine %>%
  mutate(has_fruit = str_detect(description, regex("fruit", ignore_case = TRUE)))

burgundy_with_fruit <- wine_1 %>%
  filter(province == "Burgundy" & has_fruit) %>%
  nrow()

total_with_fruit <- wine_1 %>%
  filter(has_fruit) %>%
  nrow()

pb_given_fruit <- burgundy_with_fruit / total_with_fruit
pb_given_fruit
```

# 3. Naive Bayes Algorithm

We train a naive bayes algorithm to classify a wine's province using:
1. An 80-20 train-test split.
2. Three features engineered from the description
3. 5-fold cross validation.

We report Kappa after using the model to predict provinces in the holdout sample.

```{r}
# TODO
library(tidyverse)
library(caret)
library(fastDummies)
library(thematic)
library(naivebayes) # New
library(tidytext) # New
theme_set(theme_dark())
thematic_rmd(bg = "#111", fg = "#eee", accent = "#eee")
wine <- readRDS(gzcon(url("https://cd-public.github.io/D505/dat/pinot.rds")))
names(wine)[names(wine) == 'id'] = 'ID'
wino <- wine %>% 
  mutate(year_f = as.factor(year)) %>% 
  mutate(cherry = str_detect(description,"cherry")) %>% 
  mutate(chocolate = str_detect(description,"chocolate")) %>%
  mutate(earth = str_detect(description,"earth")) %>%
  select(-description, year)

glimpse(wino)

set.seed(505)
wine_index <- createDataPartition(wino$province, p = 0.80, list = FALSE)
train <- wino[ wine_index, ]
test <- wino[-wine_index, ]

train_control <- trainControl(method = "cv", number = 5)
nb_model <- train(province ~ ., data = trainData, method = "naive_bayes", trControl = train_control)

predictions <- predict(nb_model, newdata = testData)
predictions <- factor(predictions, levels = levels(testData$province))

conf_matrix <- confusionMatrix(predictions, testData$province)
kappa_value <- conf_matrix$overall['Kappa']
kappa_value #This Kappa value is absolutely atrocious, not sure where I went awry.
```

# 4. Frequency Differences

We find the three words that most distinguish New York Pinots from all other Pinots.

```{r}
# TODO
library(tidyverse)
library(tidytext)

word_counts <- wine %>%
  unnest_tokens(word, description) %>%
  count(province, word, sort = TRUE) %>%
  group_by(province) %>%
  mutate(freq = n / sum(n)) %>%
  ungroup()

ny_freq <- word_counts %>%
  filter(province == "New York") %>%
  select(word, freq) 

other_freq <- word_counts %>%
  filter(province != "New York") %>%
  group_by(word) %>%
  summarise(freq_other = mean(freq, na.rm = TRUE)) %>%
  ungroup()

word_diff <- ny_freq %>%
  left_join(other_freq, by = "word") %>%
  mutate(diff = freq - freq_other) %>% # Math is fun, why not do it with words?
  filter(!is.na(diff)) %>%  # Attempt to remove NAs in an effort to avoid blank table
  arrange(desc(diff))

top_words_ny <- word_diff %>%
  slice_max(diff, n = 3) %>%
  select(word, diff)

print(top_words_ny)
```

# 5. Extension

Calculate the variance of the logged word-frequency distributions for each province.

```{r}
# TODO
word_freq <- wine %>%
  unnest_tokens(word, description) %>%
  count(province, word) %>%
  group_by(province) %>%
  mutate(freq = n / sum(n)) %>%
  ungroup()

word_freq <- word_freq %>%
  mutate(log_freq = log(freq + 1))

variance_by_province <- word_freq %>%
  group_by(province) %>%
  summarise(variance = var(log_freq, na.rm = TRUE))

variance_by_province

```
# After my abysmal results for questions 3 and 4, I gave my code to Chatgpt as well as the prompt and a link for the database. When yielding the floor to the supreme overlord, it had this to say:

```{r}
# TODO 
# Load required libraries
library(tidyverse)
library(caret)
library(tm)
library(e1071)

# Load dataset
wine <- readRDS(gzcon(url("https://cd-public.github.io/D505/dat/pinot.rds")))

# Ensure province is a factor
wine$province <- as.factor(wine$province)

# Preprocessing: Create a corpus from wine descriptions
corpus <- VCorpus(VectorSource(wine$description))
corpus <- tm_map(corpus, content_transformer(tolower))
corpus <- tm_map(corpus, removePunctuation)
corpus <- tm_map(corpus, removeNumbers)
corpus <- tm_map(corpus, removeWords, stopwords("en"))
corpus <- tm_map(corpus, stripWhitespace)

# Convert corpus into a Document-Term Matrix (DTM)
dtm <- DocumentTermMatrix(corpus)

# Convert DTM into a dataframe with word counts
word_freq <- as.data.frame(as.matrix(dtm))
word_freq <- cbind(province = wine$province, word_freq)

# Feature selection: Pick the top 3 most frequent words
word_sums <- colSums(word_freq[,-1])
top_words <- names(sort(word_sums, decreasing = TRUE)[1:3])

# Prepare final dataset with selected words
df <- wine %>% 
  select(province) %>%
  cbind(word_freq[, top_words])

# Ensure province is still a factor
df$province <- as.factor(df$province)

# Train-test split (80-20)
set.seed(123)
train_index <- createDataPartition(df$province, p = 0.8, list = FALSE)
train_data <- df[train_index, ]
test_data <- df[-train_index, ]

# Ensure test data has the same factor levels as training data
test_data$province <- factor(test_data$province, levels = levels(train_data$province))

# Train Naïve Bayes model with 5-fold cross-validation
train_control <- trainControl(method = "cv", number = 5)
nb_model <- train(province ~ ., data = train_data, method = "nb", trControl = train_control)

# Predict on test data
test_pred <- predict(nb_model, newdata = test_data)

# Ensure predictions are also factors with the same levels
test_pred <- factor(test_pred, levels = levels(test_data$province))

# Evaluate model performance using Kappa
conf_matrix <- confusionMatrix(test_pred, test_data$province)
kappa_value <- conf_matrix$overall["Kappa"]
print(paste("Kappa:", kappa_value))

# Find words that most distinguish New York Pinots from others
ny_pinot <- subset(df, province == "New York")
other_pinot <- subset(df, province != "New York")

# Compute word frequency ratios
ny_word_freq <- colSums(ny_pinot[,-1]) / nrow(ny_pinot)
other_word_freq <- colSums(other_pinot[,-1]) / nrow(other_pinot)

# Identify the three most distinguishing words
word_ratios <- sort(ny_word_freq / (other_word_freq + 1e-6), decreasing = TRUE)
distinguishing_words <- names(word_ratios)[1:3]
print("Three words that most distinguish New York Pinots:")
print(distinguishing_words)
```
